Random Forests, LAs

### 2024-02-15 src/la_RF_OLD.Rmd
mac

## Spatial data

@@

@@

@@

@@

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "n.London.lag", "London's website density, t-1", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.l.slag", "spatial and temporal lag of website density",
                                                             ifelse(rowname == "n.nearest.city.lag", "nearest city's website density, t-1", rowname)))))) %>% 
  column_to_rownames()

@@

## RF growth with CAST

### train the model in all data, growth

```{r}
set.seed(123)
indices <- CreateSpacetimeFolds(df %>% filter(year > 1999),                     # FILTER BY YEAR FOR GROWTH
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

start.time <- Sys.time()
model.all <- train(growth ~                                                     # REPLACE n WITH growth
                     #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
                     #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist, #n.nearest.retail.lag + dist.retail,
                   data = df %>% filter(year > 1999),                           # FILTER BY YEAR FOR GROWTH
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

# [reproducible results]
# Random Forest 
# 
# 4862 samples
#    6 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 4044, 3685, 4056, 4056, 4032, 3707, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE       Rsquared   MAE      
#   2     variance    0.2004565  0.6285613  0.1460545
#   2     extratrees  0.2001702  0.6342523  0.1480880
#   4     variance    0.2080782  0.6066952  0.1503476
#   4     extratrees  0.2065532  0.6220336  0.1527747
#   6     variance    0.2099617  0.5996347  0.1507628
#   6     extratrees  0.2084788  0.6180880  0.1544252
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 2, splitrule = extratrees and min.node.size = 5.

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "n.London.lag", "London's website density, t-1", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.l.slag", "spatial and temporal lag of website density",
                                                             ifelse(rowname == "n.nearest.city.lag", "nearest city's website density, t-1", rowname)))))) %>% 
  column_to_rownames()

path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_growth.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)
```

### train the model in a loop for all but one regions, growth

  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD!=i & year > 1999),      # FILTER BY YEAR FOR GROWTH 
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)
  # CV
  tc <- trainControl(method = "cv", #MOVED HERE< DOUBLE CHECK
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   index = indices$index,
                   savePredictions = 'final')

  model.all <- train(growth ~                                                   # REPLACE n WITH growth
                       #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                       #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
                       n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
                   data = df %>% filter(RGN21CD!=i & year > 1999),              # FILTER BY YEAR FOR GROWTH
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data
}

### Resamples for plots, growth

### Variable importance, growth

@@

### Test on one region, growth

df.predictions <- data.frame()
for (i in regions){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], 
                    df[df$RGN21CD==i & df$year > 1999,])                                # FILTER BY YEAR
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", ".on.", i), pred)

    d <- paste0("region.predict.model.", ".on.", i)
    d <- cbind(get(d), df %>% filter(RGN21CD==i & year > 1999))                         # FILTER BY YEAR
    d <- d %>% dplyr::select(RGN21CD, ladcd, year, growth, 1)                           # REPLACE n WITH growth
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(tested.on = i)
    df.predictions <- rbind(d, df.predictions)
}

# calculate metrics for every region pair 
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$growth))   # REPLACE n WITH growth

path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth.csv")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "test.on") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(RGN21NM, Rsquared) %>% 
  rename(Region = RGN21NM) %>% 
  write_csv(path.out)
  #ggplot(aes(x = 1, y = reorder(Region, Rsquared), fill = Rsquared)) +geom_tile()

>>>>

## spatial data NEW

@@

## spatial data

```{r, eval=FALSE}
# get LA for the UK
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
la <- spTransform(la, CRS("+init=epsg:4326"))

# cities
cities <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  arrange(pop) %>% tail(10) 
cities.sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)
```

@@

## Spatial panel

```{r, eval=F}
#spatial weight matrix
sf_use_s2(FALSE)
nb <- poly2nb(la, queen = T, snap = 1)
listw <- nb2listw(nb, style = "W", zero.policy = T)

test <- la
test@data <- test@data %>% left_join(df[df$year==t,], by = c("LAD21CD"="ladcd")) %>% 
  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
         year = ifelse(is.na(year), t, year))

# Load required libraries
library(spdep)
library(splm)

# Assuming you have a dataset named 'data' containing your variables: Y, W, Wc, and X

# Create spatial lag matrices
W <- mat2listw(W, style="W")
Wc <- mat2listw(Wc, style="W")

# Estimate spatial panel model
spatial_panel_model <- spml(n ~ density + London + dist + dist.retail, 
                            listw=listw,
                            data=df %>% filter(!is.na(n)), 
                            index = c("ladcd", "year"),
                            effect = "individual",
                            #method = "BFGS", # eigen
                            model= "random", 
                            lag=TRUE, 
                            na.action = na.omit) #na.fail, 
                            #zero.policy = NULL)

m.spreml <- spreml(n ~ density + London + dist + dist.retail, 
                            w=listw,
                            data=df %>% filter(!is.na(n)), 
                            index = c("ladcd", "year"),
                            effect = "individual",
                            #method = "BFGS", # eigen
                            lag=TRUE, 
                            na.action = na.omit) #na.fail, 
                            #zero.policy = NULL)

# Print summary of the model
summary(m.spreml)
```

calculate k neighbours and find out why can't have spatial FE. Hierarchical?

```{r, eval=F}
#install.packages('SDPDmod')
library(SDPDmod)

#spatial weights matrix
sf_use_s2(FALSE)
nb <- poly2nb(la, queen = T, snap = 1)
listw <- nb2listw(nb, style = "W", zero.policy = T)

#spatial weight matrix
binary_matrix <- nb2mat(nb, style = "B", zero.policy = TRUE)
class(binary_matrix)
w.binary <- rownor(binary_matrix)
isrownor(w.binary)
w.binary <- rownor(binary_matrix)
isrownor(w.binary)

#spatial weights matrix
library("sf")
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la.sf <- st_read(path.geo)
W_2n <- mOrdNbr(sf_pol = la.sf, m = 2) ## second order neighbors
W_knn <- mNearestN(distMat = gN3dist, m = 5) ## 5 nearest neighbors

#spatial weights matrix
sf_use_s2(FALSE)
la.c <- st_centroid(la.sf)
kn <- knearneigh(la.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn_binary_matrix <- nb2mat(knn, style = "W", zero.policy = TRUE)
isrownor(knn_binary_matrix)

mod2<-SDPDm(formula = n ~ density,# + London,# + dist + dist.retail, 
            #W = w.binary,
            #W = binary_matrix,
            #W = knn_binary_matrix,
            W = knn_binary_matrix,
            data=df %>% drop_na(),
            index = c("ladcd", "year"),
            model = "sar", 
            effect = "twoways",
            dynamic = T,
            LYtrans = TRUE, #it affects coefficients, need to read about it
            tlaginfo = list(ind = NULL, tl = T, stl = T),
            )
summary(mod2)

# if distances are included, effect = year. 
# Rho is positive and sig, AR is positive and sig, BUT Wn(t-1) is negative and sig
```

@@

## RF

### train on all but one regions

```{r}
# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   savePredictions = 'final')

# London region: E12000007
# training dataset
tr.df <- df %>% ungroup() %>% filter(RGN21CD!="E12000007"|
                                     RGN21CD!="E12000008")

set.seed(71)
m.test <- train(n ~ density + area + dist + dist.retail + London +
                  n.slag + year + n.nearest.city, #n.l.slag
                data = tr.df,
                trControl = tc,
                method = "rf", 
                importance = TRUE,
                na.action = na.omit)
summary(m.test)
plot(m.test)
m.test
print(m.test)

varimp_mars <- varImp(m)
plot(varimp_mars, main="Variable Importance")
```

### test on the hold out regions

Wales region is missing!!!

```{r}
pred <- predict(m, df %>% ungroup() %>% filter(RGN21CD=="E12000008"), na.action = na.pass)
length(pred)

df %>% ungroup() %>% filter(RGN21CD=="E12000008") %>% filter(if_any(everything(), is.na))

pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(RGN21CD=="E12000008"))
postResample(pred = pred$value, obs = pred$n) ## 0.99
```

## time slices

```{r}
help <- df %>% createTimeSlices(initialWindow = 5,
                        horizon = 1,
                        fixedWindow = F)

#Adding t-1 lag for n

set.seed(71)
m.test.time <- train(n ~ area + dist + dist.retail + London + #density
             n.lag + n.slag + n.l.slag + n.nearest.city + year,
           data = df %>% filter(year==1997 | year == 1998),
           trControl = tc,
           method = "rf", 
           importance = TRUE,
           na.action = na.omit)
summary(m.test.time)
plot(m.test.time)
m.test.time

varimp_mars <- varImp(m.test)
plot(varimp_mars, main="Variable Importance")

pred <- predict(m.test, df %>% ungroup() %>% filter(year==1999), na.action = na.pass)
length(pred)

pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(year==1999))
postResample(pred = pred$value, obs = pred$n) ## 0.60 
```

## RF rolling

### train the model in a loop

TODO:

- [x] read more about trainColntrol
- [ ] other variables, LA Classification, check regressions, 
- [x] N/S
- [x] reproducible
- [ ] n vs. log(n/n.lag), %>% mutate(n.lag = n.lag + 0.00001, n = n + 0.00001, way to bring S?
- [x] center or scale with preProcess?? preProc = c("center", "scale") makes no difference
- [ ] direction of relationship, or cor()
- [x] importance over time
- [x] spatial sampling
- [ ] hierarchy: ranking correlation and line graphs 
- [ ] OA

```{r}
# Train on t & t + 1

# Make the parallelisation reproducible using https://stackoverflow.com/questions/13403427/fully-reproducible-parallel-models-using-caret. 
# See also https://stackoverflow.com/questions/27944558/set-seed-parallel-random-forest-in-caret-for-reproducible-result

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   savePredictions = 'final')

# run the model for every 2 years
start.time <- Sys.time()
years <- 1997:2012

for (t in years){
  tplus1 <- t + 1
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.l.slag + n.nearest.city + year + n.lag,
                     data = df %>% filter(year==t | year==tplus1), #%>% 
                       # mutate(growth = ifelse(growth == -Inf | growth == Inf, NA, growth)) %>% 
                       # mutate(growth = ifelse(is.na(growth), 0, growth)), 
                     #%>% mutate(growth = ifelse(growth== -Inf | growth == Inf, NA, growth))
                     trControl = tc,
                     method = "ranger", #"rpart", # rf lm
                     na.action = na.omit,
                     #preProc = c("center", "scale"),
                     importance = 'impurity') # TRUE for rf
    assign(paste0("model.all",as.character(t + 1)), model.all)
}

# df %>% filter(growth == Inf) 
# log(n/n.lag)

# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
```

```{r}
# create a list of the above model outputs
fit.model.all <- mget(ls(pattern = "^model.all\\d"))

### test

# growth or n

```{r}
## Test on t + 2

# It will create: 2001 (train on 2000-2001) - 2011 (train on 2010)
# I need: 2001 (train on 2000-2001) - 2009 (train on 2008-2009)

# use the model trainned in year t to predict io of year t + 1
years <- (1998:2011)
pred.by.year.all <- data.frame(matrix(NA,nrow=374, ncol = 0))
for (t in years){
  t_plus_1 <- t + 1
  pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",t)], df[df$year==t_plus_1,])
  pred <- as.data.frame(pred)
  rownames(pred) <- c()
  pred.by.year.all <- cbind(pred.by.year.all, pred)
}

#change column names to match with total
names(pred.by.year.all) <- 1999:2012

# wide to long
pred.by.year.all <- gather(pred.by.year.all,key = year,value = predictions,1:14)

# drop 2000 and select variables form total
df.no1997_1 <- df %>% 
  filter(year!=1996 & year!=1997 & year!=1998) %>% 
  dplyr::select(ladcd, year, n) # CHANGE n WITH growth

# column bind prediction and data  
pred.by.year.all <- cbind(pred.by.year.all, df.no1997_1 %>% arrange(year, ladcd)) #DOUBLE CHECK ARRANGE

# pred.by.year to list by year
pred.by.year.all.list <- split(pred.by.year.all, pred.by.year.all$year)

# calculate metrics for every year
rf.year.all.metrics <- lapply(pred.by.year.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))  # CHANGE n WITH growth
rf.year.all.metrics %>% as.data.frame() %>%
  round(2) %>%
  kable()

# there are 2 columns named year, so I drop one
pred.by.year.all$year <- NULL

ggplot(data = pred.by.year.all,aes(x = n, y = predictions)) +
  geom_point(colour = "blue") +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
  facet_wrap(~ year,ncol = 2) +
  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
  ggtitle("Predicted vs. actual trade by year")
```

## RF spatial

df <- df %>% dplyr::select(-n.lag)

### train the model in a loop

```{r}
#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   savePredictions = 'final')

# run the model for every 2 years
start.time <- Sys.time()
regions <- df %>% distinct(RGN21CD)
regions <- as.vector(regions$RGN21CD)

for (i in regions){
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.nearest.city + year, #n.lag +  + n.l.slag
                     data = df %>% filter(RGN21CD==i), trControl = tc,
                     method = "ranger", 
                     na.action = na.omit,
                     #preProc = c("center", "scale"),
                     importance = TRUE)
  assign(paste0("model.all",as.character(i)), model.all)
}

# df %>% filter(growth == Inf) 
# log(n/n.lag)

# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
```

### test

# growth or n

```{r}
## Test on t + 2

# train on one region, test on all the rest

df.predictions <- data.frame()
for (i in regions){
  for (j in regions[!regions == i]){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", i, ".on.", j), pred)

    d <- paste0("region.predict.model.", i, ".on.", j)
    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
    d <- d[,c(5, 2, 4, 6, 1)]
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(trained.on = i)
    df.predictions <- rbind(d, df.predictions)
  }
}

df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))

# split to list by region pair
pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11

# calculate metrics for every region pair 
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))  # CHANGE n WITH growth

rf.year.all.metrics %>% 
  as.data.frame() %>%
  add_rownames(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
  rename(train.region=RGN21NM) %>% 
  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
  rename(test.region=RGN21NM) %>% 
  arrange(Rsquared)
  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()
```

### test plots

```{r}
#plots
ggplot(data = df.predictions, aes(x = n, y = predictions)) +
  geom_point(colour = "blue") +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
  facet_wrap(~ test.train, ncol = 2) +
  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
  ggtitle("Predicted vs. actual trade by year")
```

@@

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "n.London.lag", "London's web density, t-1", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.l.slag", "spatial and temporal lag of web density",
                                                             ifelse(rowname == "n.nearest.city.lag", "nearest city's web density, t-1", rowname)))))) %>% 
  column_to_rownames()

@@

#### OLD

```{r, eval=FALSE}
# train on one region, test on all the rest

df.predictions <- data.frame()
for (i in regions){
  for (j in regions[!regions == i]){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", i, ".on.", j), pred)

    d <- paste0("region.predict.model.", i, ".on.", j)
    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
    d <- d[,c(5, 2, 4, 6, 1)]
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(trained.on = i)
    df.predictions <- rbind(d, df.predictions)
  }
}

df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))

# split to list by region pair
pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11

# calculate metrics for every region pair 
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))  # CHANGE n WITH growth

path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
  rename(train.region=RGN21NM) %>% 
  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
  rename(test.region=RGN21NM) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(train.region, test.region, Rsquared) %>% 
  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
  geom_tile() +
  coord_fixed() +
  labs(y = "", x = "") + #y: test.region, x: train.region
  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
  theme(#legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust=1),
        panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white"))
ggsave(path.out)

  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()

@@
